MS Regressions

### 2021-12-02 lab_notebook/lab_notebook.Rmd
group analyses wiwth depression network

20211202
[] figure out why registration broke

>>>>

on local machine, make binary mask
```{bash}
cd /Users/eballer/BBL/msdepression/templates/harvard_depression
3dcalc -a Depression_Clust_mask_rois.nii.gz -expr 'ispositive(a)' -prefix Depression_Clust_mask_roi_binarized.nii.gz
scp -r harvard_depression eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/.
```

On cluster
```{bash}
cd /project/msdepression/templates/harvard_depression
3dresample -master AF_L.nii.gz -prefix "resampled_hcp_space_harvard_dep_mask.nii.gz" -input Depression_Clust_mask_roi_binarized.nii.gz
```

code for calc_vol_fascicles_within_depression_mask.sh 
```{bash calc_vol_fascicles_within_depression_mask}
#### calculate volume of fascicles within the depression network ####
### Pre: All streamlines must have been made in dsi studio (these are the nii files)
### Post: A depression mask resampled to template space, file <streamline_volume_within_dep_network>, containing each fascicle and its corresponding % within depression netw
ork
### Uses: Trying to figure out which fascicles are in the depression network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only l
ooking at fascicles within depression network)
### Dependencies: This script requires afni to be installed (currently using 20.1)

#set outfile
outfile="/project/msdepression/results/streamline_volume_within_dep_network.csv"

#set template directories, will be making our resampled mask from within this
template="Depression_Clust_mask_roi_binarized.nii.gz"
template_dir="/project/msdepression/templates/harvard_depression"
resampled_filename="resampled_${template}"
template_fullpath="/project/msdepression/templates/harvard_depression/${template}"

#set fascicle directory
fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
fascicle_type="association"
fascicle_for_resampling="AF_L.nii.gz"

#set home directory
homedir="/project/msdepression/scripts/"

#initiate csvs
rm -f $outfile
touch $outfile
echo "fascicle,num_voxels_total,non_zero_voxels_in_dep_map,prop_in_mask" >> $outfile

#resample Depression_Clust_amsk_roi_binarized.nii.gz to match sample fascicle - AF_L.nii.gz
if [ -f "${template_dir}/${resampled_filename}" ]; then
	rm -f ${template_dir}/${resampled_filename}
fi
3dresample -master ${fascicle_direc}/${fascicle_type}/${fascicle_for_resampling} -prefix ${template_dir}/${resampled_filename} -input ${template_fullpath} -rmode NN

#go through each fiber, get volume, and store it in csv
fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"
for fiber_bundle_type in ${fiber_bundle_type_list}; do
    echo "Fiber bundle type is " $fiber_bundle_type
    fascicle_volumes=$(ls ${fascicle_direc}/${fiber_bundle_type}/*.nii* | grep -v "dep_mask") #list all files, exclude depression ones
    for fascicle in ${fascicle_volumes}; do
	    fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.gz/$1/g')
	    num_voxels_whole_mask=$(3dmaskave -quiet -mask SELF -sum ${fascicle})

	    #remove previous mask if already made
	    if [ -f "${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz" ]; then
		    rm -f ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz
	    fi

	    3dcalc -a ${template_dir}/${resampled_filename} -b ${fascicle} -expr 'a*b' -prefix ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz
	    num_nonzero_volumes=$(3dmaskave -quiet -mask SELF -sum ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz)
	    prop_in_mask=$(echo ${num_nonzero_volumes}/${num_voxels_whole_mask} | bc -l)
	    echo "num_voxels" $num_voxels_whole_mask "num_nonzero" $num_nonzero "prop_in_mask" $prop_in_mask

	    echo "${fascicle_prefix},${num_voxels_whole_mask},${num_nonzero_volumes},${prop_in_mask}" >> $outfile
    done
done
```

Moved back to home directory
```{bash}
#from local machine
scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/streamline_volume_within_dep_network.csv .

#moved 
```

get volume in dep network per fascicle
20211202
[] figure out why registration broke

### 2021-12-02 msdepression_regressions_started_20211118.Rmd
group analyses wiwth depression network

#fascicle_mapping, returns numerical mapping as well as names of tracts
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
##########################################################
#             Step 6 - Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

@@

#visreg
sapply(fascicle_lm, visreg)
```

@@

@@

@@

@@

```
```{r depression_main_effect_mixed_model}
##### repeat above analysis but use only the first instance of each EMPI, sorted by date
df_mm <-df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]

@@

@@

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  ungroup()
##

@@

@@

>>>>

########
#fascicle volumes in depression network
########
fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep_network %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)

#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()

##########################################################
#             Step 6 - Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

@@

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

@@

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

@@

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)
#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

@@

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

@@

```{r depression_main_effect_mixed_model}
##### repeat above analysis but use only the first instance of each EMPI, sorted by date
df_mm <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]

@@

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

@@

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
#also make columns for the mean across the full depression network, and the 10% cutoff depression network
#remove rows where the total_fascicle_prop_lost is 0 - this was usually a failure of normalization
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  filter(total_fascicle_prop_lost > 0) %>%
  ungroup()

@@

dep_vs_healthy_net <- lmerTest::lmer(total_fascicle_prop_lost_indepnet ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

dep_easy_dep_net <- lm(total_fascicle_prop_lost_indepnet ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

dep_vs_healthy_net_10_percent <- lmerTest::lmer(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

dep_easy_dep_net_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

@@

#fdr corrected nondep, number of fascicles is total, 40, 5 correct - C_PHP_L, SLF1_L, UF_L, R, DRTT_L
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected dep, number of fascicles is total = 47, num correct =30
#take home - areas of worsening disease over age much more likely to be in depression network
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
{
   lmerTest::lmer(substitute(i ~ depGroupVar*PAT_AGE_AT_EXAM + (1|EMPI) , list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_agedep_lm) <- fascicle_names

fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

#fdr corrected only dep
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10%
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected nondep, number of fascicles is total, 40, 4 correct - C_PHP_L, UF_L, R, DRTT_L
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected dep, number of fascicles is total = 47, num correct = 29
#take home - areas of worsening disease over age much more likely to be in depression network
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)
```
